---
title: "Temporal patterns of trait variation"
author: "Pat Dumandan"
date: "12/01/2020"
output: html_document
---
## Overarching Question: if processes that drive population dynamics and shape community structure vary across spatial and temporal scales (e.g., effect of disturbance occurring in a small plot at year one vs effect of similar disturbance at the regional level after 10 years), then we would expect that the traits expressed by co-occurring species and individuals would also vary. But how? (i.e., in what context would we expect intra- or interspecific trait variation to be greater and how does that relate to their population and community dynamics?)

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
require(dplyr)
require(raster)
require(ggplotgui)
require(ggplot2)

Portal_rodent=read.csv("./PortalData/Rodents/Portal_rodent.csv")
Portal_plots=read.csv("./PortalData/SiteandMethods/Portal_plots.csv")
Portal_sp=read.csv("./PortalData/Rodents/Portal_rodent_species.csv")
portal1=inner_join(Portal_rodent, Portal_plots, by= c("plot", "month", "year"))
portal1$speciescode=portal1$species
portal2=inner_join(portal1, Portal_sp, by="speciescode")                  

unid_genus=c("Onychomys sp.", "Sigmodon sp.", "Chaetodipus/Peromyscus sp.", "Rodent sp.", "Dipodomys sp.")
rodent_all=portal2%>%
          filter(taxa=="Rodent", treatment=="control", !(scientificname %in% unid_genus))

```
### Q1: How do we expect intra- and interspecific trait variation to scale across different time scales?  
*calculate average coefficient of variation on a monthly, seasonal, annual, and decadal basis *  

#### **A. Monthly scale** 

#### Interspecific trait variation at different temporal scales (Interspecific CV) 
*how dispersed are the values of body size across all species across all months?*  

*calculated by taking the mean trait value for each species for each month across all years and then calculating the CV across species and month (Jung et al., 2010)*
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
month_inter=rodent_all%>%
  filter(treatment=="control", !(age=="J"),!is.na(wgt))%>%
  group_by(month, scientificname)%>%
  summarise(mean_wgt=mean(wgt))

cv(month_inter$mean_wgt) #96.64
```
#### Intraspecific trait variation at different temporal scales (Intraspecific CV) 
*how dispersed are the values of body size for each species across all months?*  
*calculated the CVs of each species for each month across all years independently and then taking the mean of these values across all 12 months (Dawson and Jonsson, 2020)*
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
month_intra=rodent_all%>%
  filter(treatment=="control", !(age=="J"),!is.na(wgt))%>%
  group_by(month, scientificname)%>%
  summarise(mean_wgt=mean(wgt))

#cv for each species for each month
mon_intra=rodent_all%>%
  filter(!is.na(wgt))%>%
  group_by(scientificname, month)%>%
  summarise(mean_wgt=mean(wgt))%>%
  mutate(cv=sd(mean_wgt)/mean(mean_wgt)*100)


mon_sp=mon_intra%>%
      group_by(scientificname)%>%
      summarise(mean_cv=mean(cv))

mean(mon_sp$mean_cv) #12.38
```

#### **B. Seasonal scale**
#### Interspecific trait variation at different temporal scales (Interspecific CV) 
*how dispersed are the values of body size across all species across all seasons?*  

*calculated by taking the mean trait value for each species for each month for each season and then calculating the CV across species and season (Jung et al., 2010)*
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
targ_s1=c(1:6)
targ_s2=c(7:12)

season1_inter=rodent_all%>%
  filter(treatment=="control", !(age=="J"),!is.na(wgt), month %in% targ_s1)%>%
  group_by(month, scientificname)%>%
  summarise(mean_wgt=mean(wgt))

cv(month_inter$mean_wgt) #96.64

```

#### Q2: At each time scale, when do we expect intra- or interspecific trait variation to be greater/more important?

