---
title: "Temporal patterns of trait variation"
author: "Pat Dumandan"
date: "12/01/2020"
output: html_document
---
## Overarching Question: if processes that drive population dynamics and shape community structure vary across spatial and temporal scales (e.g., effect of disturbance occurring in a small plot at year one vs effect of similar disturbance at the regional level after 10 years), then we would expect that the traits expressed by co-occurring species and individuals would also vary. But how? (i.e., in what context would we expect intra- or interspecific trait variation to be greater and how does that relate to their population and community dynamics?)

### Exploratory Analysis  

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
require(dplyr)
require(raster)
require(ggplotgui)
require(ggplot2)

Portal_rodent=read.csv("./PortalData/Rodents/Portal_rodent.csv")
Portal_plots=read.csv("./PortalData/SiteandMethods/Portal_plots.csv")
portal1=inner_join(Portal_rodent, Portal_plots, by= c("plot", "month", "year"))
portal2=inner_join(portal1, Portal_rodent_species, by="speciescode")                  
portal1$speciescode=portal1$species

unid_genus=c("Onychomys sp.", "Sigmodon sp.", "Chaetodipus/Peromyscus sp.", "Rodent sp.", "Dipodomys sp.")
rodent_all=portal2%>%
          filter(taxa=="Rodent", treatment=="control", !(scientificname %in% unid_genus))

unique(rodent_all$scientificname)

```
#### Q1: How do we expect intra- and interspecific trait variation to scale across different time scales?  

*calculate average coefficient of variation on a monthly, seasonal, annual, and decadal basis *
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# A. what is the average intra- and interspecic trait variation across all months?




```

#### Q2: At each time scale, when do we expect intra- or interspecific trait variation to be greater/more important?

